---
title: "FA10_KHAFAJI_RODILLAS"
author: "Rodillas"
date: "2024-11-26"
output: github_document
---

```{r}
library(tidyr)
library(dplyr)
data <- read.csv("E:/Downloads/Cholesterol_R2.csv")

long_data <- data %>%
  pivot_longer(cols = c(Before, After4weeks, After8weeks),
               names_to = "Time",
               values_to = "Cholesterol")

long_data$Time <- factor(long_data$Time, levels = c("Before", "After4weeks", "After8weeks"))
long_data$Margarine <- as.factor(long_data$Margarine)

head(long_data)
```

Assumption 4
```{r}
library(ggplot2)
boxplot(Cholesterol ~ Time, data = long_data, main = "Cholesterol Levels Over Time", xlab = "Time", 
        ylab = "Cholesterol Level", col = c("dodgerblue", "pink", "green"), outlier.col = "red")
```
There are no significant outliers.

Assumption 5
```{r}
library(dplyr)
library(e1071)
library(stats)

descriptive_measures <- long_data %>%
  group_by(Margarine, Time) %>%
  summarise(
    Valid = n(),
    Mode = paste(names(sort(table(Cholesterol), decreasing = TRUE)[1]), collapse = ", "), 
    Median = median(Cholesterol),
    Mean = mean(Cholesterol),
    `Std. Deviation` = sd(Cholesterol),
    Variance = var(Cholesterol),
    Skewness = e1071::skewness(Cholesterol),
    `Std. Error of Skewness` = sqrt((6 * Valid * (Valid - 1)) / ((Valid - 2) * (Valid + 1) * (Valid + 3))),
    Kurtosis = e1071::kurtosis(Cholesterol),
    `Std. Error of Kurtosis` = sqrt(24 / Valid),
    Minimum = min(Cholesterol),
    Maximum = max(Cholesterol),
    `25th Percentile` = quantile(Cholesterol, 0.25),
    `50th Percentile` = quantile(Cholesterol, 0.50),
    `90th Percentile` = quantile(Cholesterol, 0.90)
  )
print(descriptive_measures)
```

```{r}
normality_results <- long_data %>%
  group_by(Margarine, Time) %>%
  summarize(p_value = shapiro.test(Cholesterol)$p.value)

print(normality_results)
```


Assumption 6
```{r}
library(car)
levene_test_check <- leveneTest(Cholesterol ~ Margarine, data = long_data)
print(levene_test_check)
```



Since the p-value (<0.05), there was no homogeneity of variances of the dependent variable for all physical activity groups, as assessed by Levene’s test of homogeneity of variances, p = 0.00208.

```{r}
ggplot(long_data, aes(x = Margarine, y = Cholesterol, fill = Margarine)) + geom_boxplot() + facet_wrap(~ Time) + theme_minimal() + 
  labs(title = "Cholesterol Levels by Margarine and Time", y = "Cholesterol")
```


Because the assumption of a parametric test has been violated due to unequal variances, we tried non-parametric approaches such as Friedman's Test for Within-Subjects Factors to test whether there are significant differences between the given time conditions, with this case, before, and after 4 and 8 weeks.
```{r}
friedman_test_result <- friedman.test(Cholesterol ~ Time | ID, data = long_data)
print(friedman_test_result)
```
```{r}
ggplot(long_data, aes(x = Time, y = Cholesterol, group = ID, color = Margarine)) +geom_line() + theme_minimal() + 
  labs(title = "Cholesterol Levels Over Time", y = "Cholesterol")
```

```{r}
mann_whitney_test_b4 <- wilcox.test(Cholesterol ~ Margarine, data = long_data %>% filter(Time == "Before"))
print(mann_whitney_test_b4)

mann_whitney_after4weeks <- wilcox.test(Cholesterol ~ Margarine, data = long_data %>% filter(Time == "After4weeks"))
print("Mann-Whitney U test for 'After4weeks':")
print(mann_whitney_after4weeks)

mann_whitney_after8weeks <- wilcox.test(Cholesterol ~ Margarine, data = long_data %>% filter(Time == "After8weeks"))
print("Mann-Whitney U test for 'After8weeks':")
print(mann_whitney_after8weeks)
```

```{r}

pairwise.wilcox.test(long_data$Cholesterol, long_data$Time, paired = TRUE, p.adjust.method = "bonferroni")
```

